---
title: "Covid 19"
author: "Sarah Medina"
date: "09/27/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

In order to determine if dying is independent of sex we will use the helper file Cat Cat, graphical, numerical, and inferential methods. 


Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```
```{r eval=FALSE}
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))
```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

````{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

  
The graphical results show that from the people that are dead a little under 40 percent are women and about 70 percent are men. The men are dying at twice the rate of women. From the people that are alive about 55 percent are women and 40 percent are men. About 1/4 more women are alive then men.




  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The numerical results indicate that of the people who are alive 56 are women and 44 are men. Of the people who are dead 36 are women and 64 are men. The numerical results show that the men are dying at twice the rate of women.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The inferential results show that men are 2.3 times more likely to die than the women are.

## Conclusions

Dying is dependent of sex the graphical, numerical, and inferential results all demonstrate that men are dying at a larger rate then the women. The inferential results show that the odds ratio is 2.3. This means that the men are 2.3 more likely to die then the women this leads us to conclude that dying is dependent of sex. You are more likely to stay alive if you are a women.